\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2024-11-26 Manuscript/manuscript.tex
comments on supp
referred to supp in main text

\subsubsection{2022 National Competitions}
\label{sec:datanational}
\jy{Owen, are there names from 100-m dash?}
\of{Yes, there were four US athletes in the 100m dash that we have data on.
Marvin Bracy, Fred Kerley, Travyon Bromwell, and Christian Coleman.  All of whom
recorded their slowest times at the US Track and Field Championship.}

@@

\eds{I also think we need to comment on the men and women's data. But within row
of the figure, don't the distributions appear similar?  If that is so, would a
reviewer ask why we didn't pool the data for men and women?}
\jy{Let's put the pooled analysis in the supplement.}
\of{Please see my email}

@@

Let $Y_{ijk}$ denote the RT of observation~$k$ in heat~$j$
of year~$i$. Conditioning on a venue effect $v_i$ for year~$i$
and a heat effect $h_{i/j}$ nested within each year~$i$, the
distribution of $Y_{ijk}$ is
$\text{GG}(\mu_{ijk}, \sigma_{ijk}, \nu)$, where
\begin{align}
\log(\mu_{ijk}) &= \beta_0 + v_i , \label{eq:mu}\\
\log(\sigma_{ijk}) &= \gamma_0 + h_{i/j} , \label{eq:sigma}
\end{align}
$v_i$ is normally distributed with mean zero and
variance~$\tau_v^2$, and $h_{i/j}$ is normally distributed with mean
zero and variance~$\tau_h^2$.
The two random effects were found useful: one capturing the venue effect, which
is used to contrast years, and the second being the heat effect, where every
race was given a unique identifier with typically five to nine observations
per race. The heat effect is important as it captures the variability in the
amount of time athletes are on the starting blocks before the gun goes off.
This model can be fit with R package \texttt{gamlss}
\citep{stasinopoulos2008generalized}.

@@

@@

The fitted parameters of the GG distribution in the
GAMLSS framework in Equations~\eqref{eq:gg}--\eqref{eq:sigma} are
summarized in Table~\ref{tab:ggfit}. Results obtained from both
excluding and including 2022 data are reported. The fixed-effect
parameters include $\beta_0$, $\gamma_0$, and $\nu$, corresponding to
the intercept of the log-location, log-scale, and shape of the GG
distribution, respectively. Random effects account for variability at
the venue level $\tau_v$ on the log-scale of the $\mu$ parameter
and at the heat level $\tau_h$ on the log-scale of the $\sigma$
parameter in the density in Equation~\eqref{eq:gg}. The variance
of the venue-level random effect is smaller than the heat-level random
effect variance, suggesting that heat-level variability in the scale
parameter is substantial, though on the dispersion parameter. When the
2022 data is included, all parameter estimates remain stable except
the standard deviation of the venue-level random-effect variance,
which increases from 0.043 to 0.058 . These results
highlight that RTs are influenced by both venue and heat-level
factors, and that the inclusion of 2022 introduces greater venue-level
variability, likely due to systematic differences in RTs that year.

@@

>>>>

\subsubsection{2022 National Competitions}
\label{sec:datanational}

@@

@@

Random effects at the venue and heat levels are incorporated to the
parameters in GG distribution in~\eqref{eq:gg}.
Let $Y_{ijk}$ denote the RT of observation~$k$ in heat~$j$
of year~$i$. Conditioning on a venue effect $v_i$ for year~$i$
and a heat effect $h_{i/j}$ nested within each year~$i$, the
distribution of $Y_{ijk}$ is
$\text{GG}(\mu_{ijk}, \sigma_{ijk}, \nu)$, where
\begin{align}
\log(\mu_{ijk}) &= \beta_0 + v_i , \label{eq:mu}\\
\log(\sigma_{ijk}) &= \gamma_0 + h_{i/j} , \label{eq:sigma}
\end{align}
$v_i$ is normally distributed with mean zero and
variance~$\tau_v^2$, and $h_{i/j}$ is normally distributed with mean
zero and variance~$\tau_h^2$.
The two random effects were found useful: one 
capturing the venue effect, which is used to contrast years, and the 
second being the heat effect, where every race was given a unique 
identifier with typically five to nine observations per race.
This model can be fit with R package \texttt{gamlss} 
\citep{stasinopoulos2008generalized}. The heat effect could be added
to the model for $\mu_{ijk}$ and the venue effect could be added to 
the model for $\sigma_{ijk}$. From our comparison using the Akaike 
Information Criterion, Models~\eqref{eq:mu}--\eqref{eq:sigma} turned 
out to be preferred to more complex models or competing models.

@@

We also conducted the same analysis with men’s and women’s data
pooled, yielding similar results. Details are provided in Section~1 of
the Supplementary Material.

@@

The results reported in this subsection are from men's data only
because our investigation found significant gender difference.
The fitted parameters of the GG distribution in the
GAMLSS framework in Equations~\eqref{eq:gg}--\eqref{eq:sigma} are
summarized in Table~\ref{tab:ggfit}. Results obtained from both
excluding and including 2022 data are reported. The fixed-effect
parameters include $\beta_0$, $\gamma_0$, and $\nu$, corresponding to
the intercept of the log-location, log-scale, and shape of the GG
distribution, respectively. Random effects account for variability at
the venue level $\tau_v$ on the log-scale of the $\mu$ parameter
and at the heat level $\tau_h$ on the log-scale of the $\sigma$
parameter in the density in Equation~\eqref{eq:gg}. The variance
of the venue-level random effect is smaller than the heat-level random
effect variance, suggesting that heat-level variability in the scale
parameter is substantial, though on the dispersion parameter. When the
2022 data is included, all parameter estimates remain stable except
the standard deviation of the venue-level random-effect variance,
which increases from 0.043 to 0.058 . These results
highlight that RTs are influenced by both venue and heat-level
factors, and that the inclusion of 2022 introduces greater venue-level
variability, likely due to systematic differences in RTs that year.

@@

Results for women's data are reported in Section~2 of the
Supplementary Material.
